---
title: "DataFest Hometowns: FAKE DATA"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)

# Required packages
library(geosphere)
library(dplyr)
library(leaflet)
library(networkD3)
library(sp)
library(rworldmap)
library(plotly)

# Generat random long and lat
rand_long <- runif(n = 60, min = 0, max = 100)
rand_lat <- runif(n = 60, min = 20, max = 60)

full <- cbind(rand_long, rand_lat)
```

Column {data-width=600}
-----------------------------------------------------------------------

### Map of Hometowns

```{r}
leaflet() %>% addTiles() %>%
    addMarkers(lng = rand_long, lat = rand_lat)
```

Column {data-width=400}
-----------------------------------------------------------------------

### Close (<= 1,000 km) Hometowns Network

```{r}
distances <- vector()
for (i in 1:nrow(full)) {
    temp_indv <- full[i, ]
    
    temp_distances <- distm(full, temp_indv) %>% data.frame
    temp_distances$id <- i
    temp_distances$other_id <- 1:nrow(full)
    names(temp_distances)[1] <- 'distance'
    distances <- rbind(distances, temp_distances)
}
distances <- distances[, c(2, 3, 1)]

# Remove self-self edges
distances <- subset(distances, id != other_id)

# Keep only nodes that are within 1,000 km of each other
distances <- subset(distances, distance <= 1000000)

simpleNetwork(distances[, 1:2], zoom = TRUE)
```

### Hometown Countries

```{r}
coords2country = function(points)
{  
    # Function from: http://stackoverflow.com/a/14342127
    countriesSP <- getMap(resolution='low')

    #setting CRS directly to that from rworldmap
    pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  

    # use 'over' to get indices of the Polygons object containing each point 
    indices = over(pointsSP, countriesSP)

    # return the ADMIN names of each country
    indices$ISO3  
}

countries <- coords2country(full)

countries_count <- table(countries)
countries_count <- countries_count[countries_count > 0] 
countries_count <- data.frame(countries_count)

plot_ly(data = countries_count, x = ~countries, y = ~Freq) %>%
    layout(xaxis = list(title = ''),
           yaxis = list(title = 'Frequency'))
```

