---
title: "DataFest 2017 | Teaser: web documents with dashboards and Shiny"
author: "[Christopher Gandrud](http://christophergandrud.blogspot.com/p/research.html) & [Dustin Tingley](http://scholar.harvard.edu/dtingley)"
date: "18 January 2017"
output:
    ioslides_presentation:
        css: datafest_slides.css
        logo: img/iqss_logo_flat.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


# Dynamic and interactive documents/dashboards

## Interactive documents/dashboards

The perspectives in this session share some really important goals:

Letting people **see data** 

Letting people **explore data**

Facilitating **decisions and insights** via data

## The R Approach

[R](https://www.r-project.org/) 

<br>

Via [RStudio](www.rstudio.com):

[flexdashboard](http://rmarkdown.rstudio.com/flexdashboard/) 

[shiny](https://www.rstudio.com/products/shiny/)



# The Power of Packages

## Packages  

The Open Source R community writes packages that enable users to **do more but code less**.

Almost everything today take a couple of lines of code.

You no longer have to be a programmer to use R.

Packages are rapidly expanding R's capabilities.

## A growing community

![](img/rplot03.svg)


## Web-native visualisations

<br>

There are a growing number of R packages that make it easy to create interactive/web native visualisations.

Many of these are based on a framework called **htmlwidgets**.

## Leaflet

```{r message=FALSE}
library(leaflet); library(dplyr)

leaflet() %>% addTiles() %>% fitBounds(0, 40, 10, 50)
```

## networkD3

```{r}
library(networkD3); data(MisLinks); data(MisNodes)

forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.7, zoom = TRUE)
```

## dygraphs

```{r}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>%
    dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

## DT

```{r}
library(DT)
datatable(iris, options = list(pageLength = 5))
```



# Serving Content

## Interactive Documents

See Tools for Reproducible Data Science session run by Christopher

<br>
<br>

[RMarkdown](rmarkdown.rstudio.com)

## Dashboard Hosting Options

Many options, with free! academic licenses

[shinyapps.io](http://www.shinyapps.io/)

[RPubs](https://rpubs.com/)

[IQSS](http://www.iq.harvard.edu/) (in collaboration with [VPAL-Research](http://vpal.harvard.edu/research) ) maintains a Shiny Server for advanced use cases (Harvard Key authentication, level 3 security)

#Examples

##Galery

[Shiny Gallery](https://shiny.rstudio.com/gallery/)

##Harvard Examples

[DGSView](https://hmdc.shinyapps.io/GovDGSExample/)

Course Planning

##Shiny as GUI for a package

The [Structural Topic Model](www.structuraltopicmodel.com) is a new machine learning algorithm for analyzing text

It also has a GUI front end for people who do not want to code directly in R! (best for not huge data sets)

library(stm)

library(stmgui)

runStmGui()


# Additional considerations


## Additional considerations

99.99% of what you will see today is **free**.

Across the University much data is managed by **old spreadsheet practices**/Researchers also generating lots of data

And databases/reporting/visualization tools are **separated** from data science tools

R easily integrates with **modern databases** (SQL etc.)/has all the statistics and machine learning tools you'd want

Our students learn R, and their employers want them to know R



 
