---
title: "Flexdasboard Student Example"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
# Load required packages, install if needed.
pkgs <- c('flexdashboard', 'gsheet', 'ggmap', 'dplyr', 'leaflet', 'sp',
          'rworldmap', 'networkD3', 'geosphere', 'plotly')

pkg_missing <- pkgs[!(pkgs %in% installed.packages()[, 1])]
if (length(pkg_missing) > 0) install.packages(pkg_missing, 
                                              repos='http://cran.us.r-project.org')


library(flexdashboard)
library(gsheet)
library(ggmap)
library(dplyr)
library(leaflet)
library(sp)
library(rworldmap)
library(networkD3)
library(geosphere)
library(plotly)

# Download survey data -----
URL <- 'https://docs.google.com/spreadsheets/d/1QQkVYYdAPYjCQRO1Oupqze7Q8WULkfJ_cmDvJERumU4/edit#gid=317007960'
student_data <- gsheet2tbl(URL)

# Find hometown longitude and latitude -----
hometowns <- student_data$`What do you consider to be your "home town"?`
hometowns <- hometowns[!is.na(hometowns)]
hometown_coords <- geocode(hometowns)
hometown_coords <- subset(hometown_coords, !is.na(lon))
```

Column {data-width=650}
-----------------------------------------------------------------------

### Map of Hometowns

```{r hometown-map}
# Map student hometowns
leaflet() %>% addTiles() %>%
    addMarkers(data = hometown_coords, lng = ~lon, lat = ~lat)
```

Column {data-width=350}
-----------------------------------------------------------------------

### Close (<= 1,000 km) Hometowns Network

```{r}
# Maximum distance for relationship
max_distance <- 1000000 # in meters

# Find distances between each home town
distances <- vector()
for (i in 1:nrow(hometown_coords)) {
    temp_indv <- hometown_coords[i, ]
    
    temp_distances <- distm(hometown_coords, temp_indv) %>% data.frame
    temp_distances$id <- i
    temp_distances$other_id <- 1:nrow(hometown_coords)
    names(temp_distances)[1] <- 'distance'
    distances <- rbind(distances, temp_distances)
}
distances <- distances[, c(2, 3, 1)]

# Remove self-self edges
distances <- subset(distances, id != other_id)

# Keep only nodes that are within 1,000 km of each other
distances <- subset(distances, distance <= max_distance)

# Plot network
simpleNetwork(distances[, 1:2], zoom = TRUE)
```

### Hometown Countries

```{r}
# Function to find countries that each hometown is in


coords2country = function(points)
{  
    # Function from: http://stackoverflow.com/a/14342127
    countriesSP <- getMap(resolution='low')

    #setting CRS directly to that from rworldmap
    pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  

    # use 'over' to get indices of the Polygons object containing each point 
    indices = over(pointsSP, countriesSP)

    # return the ADMIN names of each country
    indices$ISO3  
}
countries <- coords2country(hometown_coords)

# Find country counts
countries_count <- table(countries)
countries_count <- countries_count[countries_count > 0] 
countries_count <- data.frame(countries_count)

# Plot counts
plot_ly(data = countries_count, x = ~countries, y = ~Freq) %>%
    layout(xaxis = list(title = ''),
           yaxis = list(title = 'Frequency'))
```

